---
title: "P8105 Homework 1"
author: "Kallan Roan kr3145"
date: "2025-09-12"
output: github_document
---

Load libraries
```{r}
library(moderndive)
library(tidyverse)
```

# Problem 1

Load dataset
```{r}
data("early_january_weather")
```

Description of dataset:

The `early_january_weather` dataset represents hourly meteorological data from weather station in `r unique(pull(early_january_weather, origin))` from `r month.name[unique(pull(early_january_weather, month))]` `r min(pull(early_january_weather, day))`-`r max(pull(early_january_weather, day))`, `r unique(pull(early_january_weather, year))`. 
The dataset has `r nrow(early_january_weather)` observations and `r ncol(early_january_weather)` variables. 

The variables include `r ls(early_january_weather)`. Variables `year`, `month`, `day`, and `hour` represent the time of recording. Variable `time_hour` represents the date and hour of the recording as a POSIXct date.  Variable `humid` represents the relative humidity, with an average of `r round(mean(pull(early_january_weather, humid)), digits = 2)` (Range: `r min(pull(early_january_weather, humid))`-`r max(pull(early_january_weather, humid))`). Variable `wind_dir` represents wind direction in degrees, with an average of  `r round(mean(pull(early_january_weather, wind_dir), na.rm = TRUE), digits = 2)` degrees (Range: `r min(pull(early_january_weather, wind_dir), na.rm = TRUE)`-`r max(pull(early_january_weather, wind_dir), na.rm = TRUE)` degrees). `wind_speed`, and `wind_gust` represents the wind speed and gust speed in mph, with an average of `r round(mean(pull(early_january_weather, wind_speed), digits = 2))` mph (Range: `r min(pull(early_january_weather, wind_speed))`-`r round(max(pull(early_january_weather, wind_speed)), digits = 2)` mph) and `r round(mean(pull(early_january_weather, wind_gust), na.rm = TRUE), digits = 2)` mph (Range: `r round(min(pull(early_january_weather, wind_gust), na.rm = TRUE))`-`r round(max(pull(early_january_weather, wind_gust), na.rm = TRUE))` mph) respectively. Variable `precip` represents precipitation in inches, with an average of `r mean(pull(early_january_weather, precip))` inches (Range: `r min(pull(early_january_weather, precip))`-`r max(pull(early_january_weather, precip))` inches). Variable `pressure` represents sea level pressure in millibars, with an average of `r round(mean(pull(early_january_weather, pressure), na.rm = TRUE), digits = 2)` millibars (Range: `r min(pull(early_january_weather, pressure), na.rm = TRUE)`-`r max(pull(early_january_weather, pressure), na.rm = TRUE)` millibars). Variable `visib` represents visibility in miles, with an average of `r round(mean(pull(early_january_weather, visib)), digits = 2)` miles (Range: `r min(pull(early_january_weather, visib))`-`r max(pull(early_january_weather, visib))` miles). Variable `dewp` represents the dewpoint in F, with an average of `r round(mean(pull(early_january_weather, dewp)), digits = 2)` F (Range: `r min(pull(early_january_weather, dewp))`-`r max(pull(early_january_weather, dewp))` F). Variable `temp` represents the temperature in F. The mean temperature is `r round(mean(pull(early_january_weather, temp)), digits = 2)` F  (Range: `r min(pull(early_january_weather, temp))`-`r max(pull(early_january_weather, temp))` F).

Scatterplot of temp (y) vs time_hour (x)
```{r}
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) + 
  geom_point()
```

Although the temperature fluctuates within each day, on average, as the date and time increases, the temperature increases. The temperature decreased dramatically after January 14. 
While the humidity started out low on January 1st, the humidity slightly increased starting January 6 on average. However, there is a huge increase in both temperature and humidity starting January 12. 

Save scatterplot to project directory
```{r}
ggsave("prob_1scatterplot.pdf", height = 4, width = 6)
```

# Problem 2

Create a dataframe
```{r}
prob2_df <-
  tibble(
    numeric_samp = rnorm(10),
    greater_0 = numeric_samp > 0,
    vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    vec_factor = factor(rep(c("a", "b", "c"), length.out = 10))
  )
```

Mean of numeric_samp
```{r}
mean(pull(prob2_df, numeric_samp))
```

Mean of greater_0
```{r}
mean(pull(prob2_df, greater_0))
```

Mean of vec_char
```{r}
mean(pull(prob2_df, vec_char))
```

Mean of vec_factor
```{r}
mean(pull(prob2_df, vec_factor))
```

The random sample of size 10 from a standard normal distribution produced a mean. The logical vector was also able to produce a mean, but the mean doesn't make sense because of the bolean values (TRUE and FALSE). The character vector and factor vector was not able to produce a value for a mean.

Convert greater_0, vec_char, and vec_factor to numeric
```{r results = "hide"}
as.numeric(pull(prob2_df, greater_0))
as.numeric(pull(prob2_df, vec_char))
as.numeric(pull(prob2_df, vec_factor))
```

as.numeric function converts previous logical, character, and factor variables into numeric variables. 
When as.numeric function is applied to the logical vector, it is able to successfully convert the values in the vector into a numeric values. This occurs because the values in the logical vector (0 and 1) are numbers that represent TRUE and FALSE. However, the values of 0 and 1 represent boolean values and don't hold additional meaning, so it does not make sense to take the mean of the logical vector, even if it's converted to a numeric vector and error message doesn't show when calculating the mean.

When as.numeric function is applied to the character vector, an error message will show. This is because the values in the vector resemble a character, not a number. You cannot assign numbers to characters. This does make sense when I am trying to take the mean, since you cannot take a mean that are not numbers. 

When as.numeric function is applied to the factor vector, it is able to successfully convert the values in the vector into a numeric values. This occurs because the values in the factor vector (1, 2, and 3) are actual numbers. However, since the represent the factor levels within the vector, they do not hold additional meaning. Therefore, calculating the mean of this variable does not make sense, even if is converted to a numeric vector and an error message doesn't show when calculating the mean.


?????






