---
title: "P8105 Homework 1"
author: "Kallan Roan kr3145"
date: "2025-09-12"
output: github_document
---

Load libraries
```{r}
library(moderndive)
library(tidyverse)
```

# Problem 1

Load dataset
```{r}
data("early_january_weather")
```

Description of dataset:

The dataset has `r nrow(early_january_weather)` observations and `r ncol(early_january_weather)` variables. The variables in the dataset include `r ls(early_january_weather)`. 

The data represents hourly meterological data from weather station in `r unique(early_january_weather$origin)` from `r month.name[early_january_weather$month[1]]` `r min(early_january_weather$day)` to `r month.name[early_january_weather$month[1]]` `r max(early_january_weather$day)`, `r unique(early_january_weather$year)`.

The average humidity is `r round(mean(early_january_weather$humid), digits = 2)` with a range of between `r min(early_january_weather$humid)` and `r max(early_january_weather$humid)`. The average 
precipitation is `r mean(early_january_weather$precip)` inches with a range between `r min(early_january_weather$precip)` inches and `r max(early_january_weather$precip)` inches. The average dewpoint is `r round(mean(early_january_weather$dewp), digits = 2)` F with a range of between `r min(early_january_weather$dewp)` F and `r max(early_january_weather$dewp)` F. The mean temperature is `r round(mean(early_january_weather$temp), digits = 2)` F with a range between `r min(early_january_weather$temp)` F and `r max(early_january_weather$temp)` F.

Scatterplot of temp (y) vs time_hour (x)
```{r}
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) + geom_point()
```

Although the temperature does fluctuate within each day, on average, as the date and time increases, the temperature increases. The temperature decreased dramatically after January 14. 
While the humidity started out low on January 1st, the humidity slightly increased starting January 6 on average. However, there is a huge increase in both temperature and humidity starting January 12. 

Save scatterplot to project directory
```{r}
ggsave("prob_1scatterplot.pdf", height = 4, width = 6)
```

