---
title: "P8105 Homework 1"
author: "Kallan Roan kr3145"
date: "2025-09-12"
output: github_document
---

Load libraries
```{r}
library(moderndive)
library(tidyverse)
```

# Problem 1

Load dataset
```{r}
data("early_january_weather")
```

Description of dataset:

The dataset has `r nrow(early_january_weather)` observations and `r ncol(early_january_weather)` variables. The variables in the dataset include `r ls(early_january_weather)`. 

The data represents hourly meterological data from weather station in `r unique(early_january_weather$origin)` from `r month.name[early_january_weather$month[1]]` `r min(early_january_weather$day)` to `r month.name[early_january_weather$month[1]]` `r max(early_january_weather$day)`, `r unique(early_january_weather$year)`.

The average humidity is `r round(mean(early_january_weather$humid), digits = 2)` with a range of between `r min(early_january_weather$humid)` and `r max(early_january_weather$humid)`. The average 
precipitation is `r mean(early_january_weather$precip)` inches with a range between `r min(early_january_weather$precip)` inches and `r max(early_january_weather$precip)` inches. The average dewpoint is `r round(mean(early_january_weather$dewp), digits = 2)` F with a range of between `r min(early_january_weather$dewp)` F and `r max(early_january_weather$dewp)` F. The mean temperature is `r round(mean(early_january_weather$temp), digits = 2)` F with a range between `r min(early_january_weather$temp)` F and `r max(early_january_weather$temp)` F.


